---
title: "PT ACE2 KO Merge and Integration "
author: "Jackie Emathinger Xiao-Tong Su"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: 
      collapsed: false
      smooth_scroll: false
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=T, error=FALSE, message=FALSE, warning=FALSE}

library("dplyr")
library("Seurat")
library("knitr")
library("ggplot2")
library("BiocManager")
library("here")
BiocManager::install("EnhancedVolcano")
library("EnhancedVolcano") #volcano plot
install.packages('DESeq2') #for DEG
library("DESeq2")
library("tidyverse") #tidy up data

if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")} # for color brewer
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")} # for data normalization
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")} # for data normalization, sctransform
if (!require("cowplot")) {install.packages("cowplot"); require("cowplot")} # for figure layout
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")} # for figure patching
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")} # to save .xlsx files
set.seed(12345)

```

# Analyze all cells


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
#Sample	Mouse ID	Genotype	Treatment	Enrichment	Cohort	Sample Processor
#1	2584	Cre -	Saline	n/a	1	JWN
#2	2580	Cre +	AngII	n/a	1	JWN
#3	2583	Cre +	Saline	n/a	1	JME
#4	2743	Cre -	AngII	n/a	1	JME
#5	2582	Cre +	Saline	n/a	2	JWN
#6	2701	Cre -	AngII	n/a	2	JWN
#7	2744	Cre -	Saline	n/a	2	JME
#8	2581	Cre +	AngII	n/a	2	JME

ctrl <- readRDS(here("Temp Outputs", "2584.rds"))
ctrl 
#An object of class Seurat 
#32272 features across 10917 samples within 1 assay 
#Active assay: RNA (32272 features, 1988 variable features)
# 2 dimensional reductions calculated: pca, umap

ctrl2<- readRDS(here("Temp Outputs", "2743.rds"))
ctrl2
# An object of class Seurat 
# 32272 features across 12450 samples within 1 assay 
# Active assay: RNA (32272 features, 1992 variable features)
# 2 dimensional reductions calculated: pca, umap

ctrl3<- readRDS(here("Temp Outputs", "2701.rds"))
ctrl3 
# An object of class Seurat 
# 32272 features across 8541 samples within 1 assay 
# Active assay: RNA (32272 features, 1991 variable features)
#  2 dimensional reductions calculated: pca, umap

ctrl4 <-readRDS(here("Temp Outputs", "2744.rds"))
ctrl4
#An object of class Seurat 
# 32272 features across 8012 samples within 1 assay 
# Active assay: RNA (32272 features, 1994 variable features)
# 2 dimensional reductions calculated: pca, umap

exp <- readRDS(here("Temp Outputs", "2580.rds"))
exp
# An object of class Seurat 
# 32272 features across 10412 samples within 1 assay 
# Active assay: RNA (32272 features, 1988 variable features)
# 2 dimensional reductions calculated: pca, umap

exp2<-readRDS(here("Temp Outputs", "2583.rds"))
exp2 

#An object of class Seurat 
#32272 features across 13413 samples within 1 assay 
#Active assay: RNA (32272 features, 1989 variable features)
# 2 dimensional reductions calculated: pca, umap

exp3<-readRDS(here("Temp Outputs", "2582.rds"))
exp3

#An object of class Seurat 
#32272 features across 11537 samples within 1 assay 
#Active assay: RNA (32272 features, 1989 variable features)
# 2 dimensional reductions calculated: pca, umap

exp4<-readRDS(here("Temp Outputs", "2581.rds")) 
exp4 
#An object of class Seurat 
#32272 features across 9248 samples within 1 assay 
#Active assay: RNA (32272 features, 1995 variable features)
# 2 dimensional reductions calculated: pca, umap


#1	2584	Cre -	Saline	n/a	1	JWN
#2	2580	Cre +	AngII	n/a	1	JWN
#3	2583	Cre +	Saline	n/a	1	JME
#4	2743	Cre -	AngII	n/a	1	JME
#5	2582	Cre +	Saline	n/a	2	JWN
#6	2701	Cre -	AngII	n/a	2	JWN
#7	2744	Cre -	Saline	n/a	2	JME
#8	2581	Cre +	AngII	n/a	2	JME
ctrl <- AddMetaData(object = ctrl, metadata = "Saline", col.name = "Treatment")
ctrl2<- AddMetaData(object=ctrl2, metadata="AngII", col.name="Treatment")
ctrl3<- AddMetaData(object =ctrl3, metadata = "AngII", col.name="Treatment")
ctrl4<-AddMetaData(object = ctrl4, metadata= "Saline", col.name= "Treatment")

exp <- AddMetaData(object = exp, metadata = "AngII", col.name = "Treatment")
exp2<-AddMetaData(object=exp2, metadata="Saline", col.name="Treatment")
exp3<-AddMetaData(object=exp3, metadata ="Saline", col.name = "Treatment")
exp4<-AddMetaData(object=exp4, metadata= "AngII", col.name="Treatment")

ctrl <- AddMetaData(object = ctrl, metadata = "2584", col.name = "Sample_ID")
ctrl2<- AddMetaData(object=ctrl2, metadata="2743", col.name="Sample_ID")
ctrl3 <- AddMetaData(object = ctrl3, metadata = "2701", col.name = "Sample_ID")
ctrl4<- AddMetaData(object=ctrl4, metadata="2744", col.name="Sample_ID")

exp <- AddMetaData(object = exp, metadata = "2580", col.name = "Sample_ID")
exp2<-AddMetaData(object=exp2, metadata="2583", col.name="Sample_ID")
exp3 <- AddMetaData(object = exp3, metadata = "2582", col.name = "Sample_ID")
exp4<-AddMetaData(object=exp4, metadata="2581", col.name="Sample_ID")

ctrl <- AddMetaData(object = ctrl, metadata = "First", col.name = "Cohort")
ctrl2<- AddMetaData(object=ctrl2, metadata="First", col.name="Cohort")
ctrl3 <- AddMetaData(object = ctrl3, metadata = "Second", col.name = "Cohort")
ctrl4<- AddMetaData(object=ctrl4, metadata="Second", col.name="Cohort")

exp <- AddMetaData(object = exp, metadata = "First", col.name = "Cohort")
exp2<-AddMetaData(object=exp2, metadata="First", col.name="Cohort")
exp3 <- AddMetaData(object = exp3, metadata = "Second", col.name = "Cohort")
exp4<-AddMetaData(object=exp4, metadata="Second", col.name="Cohort")

ctrl <- AddMetaData(object = ctrl, metadata = "JWN", col.name = "Sample_Processor")
ctrl2<- AddMetaData(object=ctrl2, metadata="JME", col.name="Sample_Processor")
ctrl3 <- AddMetaData(object = ctrl3, metadata = "JWN", col.name = "Sample_Processor")
ctrl4<- AddMetaData(object=ctrl4, metadata="JME", col.name="Sample_Processor")

exp <- AddMetaData(object = exp, metadata = "JWN", col.name = "Sample_Processor")
exp2<-AddMetaData(object=exp2, metadata="JME", col.name="Sample_Processor")
exp3 <- AddMetaData(object = exp3, metadata = "JWN", col.name = "Sample_Processor")
exp4<-AddMetaData(object=exp4, metadata="JME", col.name="Sample_Processor")
```

### Merge datasets 

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
all_list<-list(ctrl, ctrl2, ctrl3, ctrl4, exp, exp2, exp3, exp4)
SO_1 <- merge(ctrl, y=c(ctrl2, ctrl3, ctrl4, exp, exp2, exp3, exp4), add.cell.ids = c("ctrl", "ctrl2", "ctrl3", "ctrl4", "exp", "exp2", "exp3", "exp4"), project = "PT ACE2 KO")
SO_1
#An object of class Seurat 
#32272 features across 84530 samples within 1 assay 
#Active assay: RNA (32272 features, 0 variable features)

SO_1@meta.data <- SO_1@meta.data %>% 
  mutate(orig.ident = dplyr::case_when(
    orig.ident == "SeuratProject" ~ "PT ACE2 KO"))

```

### Sanity check

```{r sanity, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

head(colnames(SO_1))
tail(colnames(SO_1))
unique(sapply(X = strsplit(colnames(SO_1), split = "_"), FUN = "[", 1))
table(SO_1$Treatment)
head(SO_1@meta.data)

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Treatment")

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Genotype")

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Cohort")

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Sample_Processor")

# Remove all nk/kd files from the environment to save on RAM
rm(ctrl, exp, ctrl2, exp2, ctrl3, exp3, ctrl4, exp4)

```

### Filter low quality cells

```{r merge, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_1 <- subset(SO_1, subset = nFeature_RNA > 500
                  & nFeature_RNA < 4000 
                  & nCount_RNA < 10000) 

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Treatment")

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Genotype")

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Sample_ID")

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Cohort")

VlnPlot(SO_1, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        pt.size = 0,
        group.by = "Sample_Processor")

SO_1 

# 1987 nuclei were filtered out

#An object of class Seurat 
#32272 features across 82543 samples within 1 assay 
#Active assay: RNA (32272 features, 0 variable features)

```

## Normalize and integrate data
### Normalize data with sctransform

```{r sctransform, echo=T, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results=F}

#Note that this single command replaces NormalizeData(), ScaleData(), and FindVariableFeatures().
#https://satijalab.org/seurat/articles/sctransform_vignette.html
#The latest version of sctransform also supports using glmGamPoi package which substantially improves the speed of the learning procedure. It can be invoked by specifying method="glmGamPoi".

SO_1 <- SCTransform(SO_1, method = "glmGamPoi", vars.to.regress = c("nCount_RNA"), verbose = TRUE)

```

### Principal Component Analysis

```{r pca, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_1 <- RunPCA(SO_1, verbose = TRUE)

ElbowPlot(SO_1, ndims = 50)

#determine percent of variation assocaited with each PC
pct <- SO_1[["pca"]]@stdev / sum(SO_1[["pca"]]@stdev) * 100
#Calculate cumulative percents for each PC
cumu <- cumsum(pct)
# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]

co1

# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

# last point where change of % of variation is more than 0.1%.
co2
# Minimum of the two calculation
pcs <- min(co1, co2)

pcs

# Create a dataframe with values
plot_df <- data.frame(pct = pct, 
           cumu = cumu, 
           rank = 1:length(pct))

# Elbow plot to visualize 
  ggplot(plot_df, aes(cumu, pct, label = rank, color = rank > pcs)) + 
  geom_text() + 
  geom_vline(xintercept = 90, color = "grey") + 
  geom_hline(yintercept = min(pct[pct > 5]), color = "grey") +
  theme_bw()
```

### Integrate data with FindIntegrationAnchors

```{r integration, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results=F}

#FindIntegrationAnchors post-scTransform. Main Differences --

# - Normalize datasets individually by `SCTransform()`, instead of `NormalizeData()` prior to integration.  
# - Use 3,000 or more features for analysis downstream of `sctransform`.  
# - Run the `PrepSCTIntegration()` function prior to identifying anchors.  
# - When running `FindIntegrationAnchors()`, and `IntegrateData()`, set the normalization.method parameter to the value `SCT`.  
# - When running sctransform-based workflows, including integration, do not run the `ScaleData()` function.

SO_1.list <- SplitObject(SO_1,
                              split.by = "Sample_ID")

SO_1.list <- lapply(X = SO_1.list,
                         FUN = SCTransform)

features <- SelectIntegrationFeatures(object.list = SO_1.list,
                                      nfeatures = 3000)

SO_1.list <- PrepSCTIntegration(object.list = SO_1.list,
                                     anchor.features = features)
saveRDS(SO_1, here("Temp Outputs", "SO_1temp"))

anchors <- FindIntegrationAnchors(object.list = SO_1.list,
                                  reduction = "rpca",
                                  normalization.method = "SCT",
                                  anchor.features = features)



SO_1 <- IntegrateData(anchorset = anchors,
                           normalization.method = "SCT")
rm(anchors, SO_1.list)
# Default assay automatically changes to “integrated” after 'IntegrateData'

SO_1 <- RunPCA(SO_1, verbose = FALSE)


SO_2 <- RunUMAP(object=SO_1, dims = 1:50)
SO_2 <- RunUMAP(SO_1, dims = 1:50)

SO_2 <- FindNeighbors(SO_2, reduction = "pca", dims = 1:50)

SO <- FindClusters(SO_2, resolution = 1)
DefaultAssay(SO) <- "SCT"

rm(SO_1, SO_2)

```

### Check batch effects

```{r split, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=5, fig.width=16}

f1 <- DimPlot(SO,
              reduction = "umap",
              label = TRUE,
              pt.size = 0,
              label.size = 6) +
  NoLegend()
#check batch effects by treatment 

SO$Treatment <- factor(x = SO$Treatment, levels = c("Veh", "AngII")) 
f2 <- DimPlot(SO, reduction = "umap", split.by = "Treatment", ncol = 2)

ggdraw() +
  draw_plot(f1, x = 0, y = 0, width = .4, height = 1) +
  draw_plot(f2, x = .4, y = 0, width = .6, height = 1)

#check batch effect by sample
SO$Sample_ID <- factor(x = SO$Sample_ID, levels = c("951", "884", "974","883")) 
f3 <- DimPlot(SO, reduction = "umap", split.by = "Sample_ID", ncol = 2)

ggdraw() +
  draw_plot(f1, x = 0, y = 0, width = .4, height = 1) +
  draw_plot(f3, x = .4, y = 0, width = .6, height = 1)

```

## Verification of Cluster Identity

```{r QC, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=5, fig.width=16}

f4 <- VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        fill.by = "ident",
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0) +
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 0, hjust = .5),
        axis.title.x = element_blank()) +
  stat_summary(fun = median,
        geom = "crossbar",
        width = 0.3,
        size = 0.2,
        position = position_dodge(width = 0.5))

ggdraw() +
  draw_plot(f1, x = 0, y = 0, width = .3, height = 1) +
  draw_plot(f4, x = .3, y = 0, width = .7, height = 1)

FeaturePlot(SO,
             features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
             cols = c("lightgrey", "royal blue"),
             ncol = 3)

```

### Identify each cluster

```{r vlnplot, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=5, fig.width=16}

markers.to.plot1 <- c(
                      "Slc5a12",      # PT-S1
                      "Lrp2",         # PT
                      "Flt1",         # Endo
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1
                      "Pdgfrb",       # Peri
                      "Slc8a1",       # DCT2, CNT
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Aqp2",         # PC
                      "Nphs1",        # Podo
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Ptprc",        # Immune
                      "Nos1",         # MD

                      "Top2a"        # Proliferation
                      )

f5 <- VlnPlot(SO, 
        features = markers.to.plot1,
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0,
        fill.by = "ident") +
  theme(legend.position = "none", 
              plot.title = element_text(hjust = 0.5),
              axis.text.x = element_text(angle = 0, hjust = .5),
              axis.title.x = element_blank()) + 
  stat_summary(fun = median,
               geom = "crossbar",
               width = 0.3,
               size = 0.1,
               position = position_dodge(width = 0.5))

f5.2 <- DotPlot(SO,
                features = markers.to.plot1,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5)+
  coord_flip()

ggdraw() +
  draw_plot(f1, x = 0, y = 0, width = .3, height = 1) +
  draw_plot(f5, x = .3, y = 0, width = .7, height = 1)

f5.2

```

### Feature plot

```{r featureplot, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 10, fig.width = 10}

f6 <- FeaturePlot(SO,
             features = markers.to.plot1,
             cols = c("lightgrey", "royal blue"),
             combine = F)

for(i in 1:length(f6)) {
  f6[[i]] <- f6[[i]] + NoLegend() + NoAxes()
}
cowplot::plot_grid(plotlist = f6)

```

### Top five genes in each cluster

```{r top5, echo=TRUE, error=FALSE, fig.align='center', fig.height=28, fig.width=10, message=FALSE, warning=FALSE}

DefaultAssay(SO) <-"RNA"

ALL.markers <- FindAllMarkers(SO, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
saveRDS(ALL.markers, "top5_all.rda")

top5 <- ALL.markers %>% group_by(cluster) %>% top_n(5, avg_log2FC)

top5

```

## Name each cluster based on their identities

```{r rename, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO@meta.data <- SO@meta.data %>% mutate(subclass.l2 = dplyr::case_when(
  seurat_clusters == 0 ~ "PT-S1",
  seurat_clusters == 1 ~ "EC1",
  seurat_clusters == 2 ~ "PT-(S2-S3)",
  seurat_clusters == 3 ~ "PT-S1",
  seurat_clusters == 4 ~ "PT-S2",
  seurat_clusters == 5 ~ "TAL",
  seurat_clusters == 6 ~ "TAL",
  seurat_clusters == 7 ~ "DCT",
  seurat_clusters == 8 ~ "Fibroblast",
  seurat_clusters == 9 ~ "PT-S3",
  seurat_clusters == 10 ~ "DCT-CNT",
  seurat_clusters == 11 ~ "EC1",
  seurat_clusters == 12 ~ "X-a",
  seurat_clusters == 13 ~ "PT-S2",
  seurat_clusters == 14 ~ "PC",
  seurat_clusters == 15 ~ "Podocyte",
  seurat_clusters == 16 ~ "PC-1",
  seurat_clusters == 17 ~ "X-b",
  seurat_clusters == 18 ~ "PC-1",
  seurat_clusters == 19 ~ "DTL-ATL?",
  seurat_clusters == 20 ~ "IC-A",
  seurat_clusters == 21 ~ "Fibroblast",
  seurat_clusters == 22 ~ "IC-B",
  seurat_clusters == 23 ~ "Immune-Macrophage",
  seurat_clusters == 24 ~ "X-c",
  seurat_clusters == 25 ~ "EC-2",
  seurat_clusters == 26 ~ "DCT-CNT",
  seurat_clusters == 27 ~ "Uro",
  seurat_clusters == 28 ~ "Pericyte",
  seurat_clusters == 29 ~ "New PT-S2",
  seurat_clusters == 30 ~ "X-c",
  seurat_clusters == 31 ~ "PEC",
  seurat_clusters == 32 ~ "DTL-PEC",
  seurat_clusters == 33 ~ "X-d",
  seurat_clusters == 34 ~ "TAL-MD",
  seurat_clusters == 35 ~ "Pericyte",
  seurat_clusters == 36 ~ "Fibroblast",
  seurat_clusters == 37 ~ "TAL, DCT1",
  seurat_clusters == 38 ~ "PT-S3",
  seurat_clusters == 39 ~ "X-e",
  seurat_clusters == 40 ~ "Doublet",
    ))

#subclas level 1 is the most complex
#subclass level 1 is a bit broader and also defines doublets based on "co-expression" of two or more genes that don't really go together in nature

SO@meta.data <- SO@meta.data %>% mutate(subclass.l1 = dplyr::case_when(
 seurat_clusters == 0 ~ "PT-S1",
  seurat_clusters == 1 ~ "Endo",
  seurat_clusters == 2 ~ "PT-(S2-S3)",
  seurat_clusters == 3 ~ "PT-S1",
  seurat_clusters == 4 ~ "PT-(S2-S3)",
  seurat_clusters == 5 ~ "TAL",
  seurat_clusters == 6 ~ "TAL",
  seurat_clusters == 7 ~ "DCT",
  seurat_clusters == 8 ~ "Fibroblast",
  seurat_clusters == 9 ~ "PT-(S2-S3)",
  seurat_clusters == 10 ~ "DCT-CNT",
  seurat_clusters == 11 ~ "Endo",
  seurat_clusters == 12 ~ "X",
  seurat_clusters == 13 ~ "PT-(S2-S3)",
  seurat_clusters == 14 ~ "PC",
  seurat_clusters == 15 ~ "Podocyte",
  seurat_clusters == 16 ~ "PC",
  seurat_clusters == 17 ~ "X",
  seurat_clusters == 18 ~ "PC",
  seurat_clusters == 19 ~ "Doublet",
  seurat_clusters == 20 ~ "IC-A",
  seurat_clusters == 21 ~ "Fibroblast",
  seurat_clusters == 22 ~ "IC-B",
  seurat_clusters == 23 ~ "Immune",
  seurat_clusters == 24 ~ "X",
  seurat_clusters == 25 ~ "Endo",
  seurat_clusters == 26 ~ "DCT-CNT",
  seurat_clusters == 27 ~ "Uro",
  seurat_clusters == 28 ~ "Pericyte",
  seurat_clusters == 29 ~ "PT-(S2-S3)",
  seurat_clusters == 30 ~ "X",
  seurat_clusters == 31 ~ "PEC",
  seurat_clusters == 32 ~ "Doublet",
  seurat_clusters == 33 ~ "X",
  seurat_clusters == 34 ~ "Macula Densa",
  seurat_clusters == 35 ~ "Pericyte",
  seurat_clusters == 36 ~ "Fibroblast",
  seurat_clusters == 37 ~ "Doublet",
  seurat_clusters == 38 ~ "PT-(s2-S3)",
  seurat_clusters == 39 ~ "X",
  seurat_clusters == 40 ~ "Doublet",
    ))

#Finally, this is the most broad class which allows you to look at the big picture!
SO@meta.data <- SO@meta.data %>% mutate(class = dplyr::case_when(
  seurat_clusters == 0 ~ "PT",
  seurat_clusters == 1 ~ "Endo",
  seurat_clusters == 2 ~ "PT",
  seurat_clusters == 3 ~ "PT",
  seurat_clusters == 4 ~ "PT",
  seurat_clusters == 5 ~ "TAL",
  seurat_clusters == 6 ~ "TAL",
  seurat_clusters == 7 ~ "DCT",
  seurat_clusters == 8 ~ "Fibroblast",
  seurat_clusters == 9 ~ "PT",
  seurat_clusters == 10 ~ "CNT",
  seurat_clusters == 11 ~ "Endo",
  seurat_clusters == 12 ~ "X",
  seurat_clusters == 13 ~ "PT",
  seurat_clusters == 14 ~ "PC",
  seurat_clusters == 15 ~ "Podocyte",
  seurat_clusters == 16 ~ "PC",
  seurat_clusters == 17 ~ "X",
  seurat_clusters == 18 ~ "PC",
  seurat_clusters == 19 ~ "Doublet",
  seurat_clusters == 20 ~ "IC-A",
  seurat_clusters == 21 ~ "Fibroblast",
  seurat_clusters == 22 ~ "IC-B",
  seurat_clusters == 23 ~ "Immune",
  seurat_clusters == 24 ~ "X",
  seurat_clusters == 25 ~ "Endo",
  seurat_clusters == 26 ~ "CNT",
  seurat_clusters == 27 ~ "Uro",
  seurat_clusters == 28 ~ "Pericyte",
  seurat_clusters == 29 ~ "PT",
  seurat_clusters == 30 ~ "X",
  seurat_clusters == 31 ~ "PEC",
  seurat_clusters == 32 ~ "Doublet",
  seurat_clusters == 33 ~ "X",
  seurat_clusters == 34 ~ "Macula Densa",
  seurat_clusters == 35 ~ "Pericyte",
  seurat_clusters == 36 ~ "Fibroblast",
  seurat_clusters == 37 ~ "Doublet",
  seurat_clusters == 38 ~ "PT",
  seurat_clusters == 39 ~ "X",
  seurat_clusters == 40 ~ "Doublet",
    ))

saveRDS(SO, here("Temp Outputs", "allcombined.rds"))

SO <- readRDS(here("Temp Outputs", "allcombined.rds"))

```

## Clusters - class

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

head(SO@meta.data)

markers.to.plot2 <- c("Slc5a12", "Lrp2", "Pdgfrb", "Nphs1", "Flt1", "Slc12a1", "Nos1", "Slc12a3", "Slc8a1", "Aqp2", "Slc4a1", "Slc26a4")

jwnlist <- c("Pdgfrb", "Lrp2")


levels(Idents(SO))


class_levels <- c("PT",	"Endo",	"TAL",	"DCT",	"Fibroblast",	"Macula Densa",	"Podocyte",	"DCT-CNT",	"PC",	"IC-A",	"IC-B", "Immune", "Uro", "PEC", "Pericyte", "Doublet",	"X")
SO@meta.data$class <- factor(x = SO@meta.data$class, levels = class_levels)
Idents(SO) <- "class"

DimPlot(SO, 
        reduction = "umap", 
        label = TRUE,
        repel = TRUE,
        pt.size = 0.5)

VlnPlot(SO, features= "Lrp2")
SO



SO2 <- ScaleData(object = SO)


pbmc <- NormalizeData(object = pbmc)
pbmc <- FindVariableFeatures(object = pbmc)
pbmc <- ScaleData(object = pbmc)


DefaultAssay(SO) <-"RNA"

DefaultAssay(SO) <-"integrated"

DefaultAssay(SO) <-"SCT"

Seurat::DotPlot(SO, features = markers.to.plot2)

head(SO@meta.data)

VlnPlot(SO, "Pdgfrb")
       
FeaturePlot(SO, features = markers.to.plot2)


DefaultAssay(SO) <-"integrated"

Seurat::DotPlot(SO,
        features = markers.to.plot2,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = .8),
        plot.title = element_text(hjust = 0.8)) +
  coord_flip()


Idents(SO)

DotPlot(SO,
        features =  markers.to.plot2,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = .8),
        plot.title = element_text(hjust = 0.8)) +
  coord_flip()

VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0,
        fill.by = "ident") +
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) +
  stat_summary(fun = median,
        geom = "crossbar",
        width = 0.3,
        size = 0.2,
        position = position_dodge(width = 0.5))

```

## Clusters - subclass level1

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=5, fig.width=7}
Idents(SO) <- "subclass.l1"
sub.l1_levels <- c("PT-S1",	"PT-S2",	"Pericyte",	"Podocyte",	"Endo",	"TAL-1",	"TAL-2",	"Macula Densa",	"DCT1",	"DCT2",	"CNT-1",	"CNT-2",	"PC",	"IC-A",	"IC-B",	"Doublet",	"X")
Idents(SO) <- factor(x = Idents(SO), levels = sub.l1_levels)



DimPlot(SO, 
        reduction = "umap", 
        label = TRUE,
        repel = TRUE,
        pt.size = 0.5) +
  ggtitle("Subclass level1")

DotPlot(SO,
        features = markers.to.plot2,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.8)) +
  coord_flip() +
  ggtitle("Subclass level1")

VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0,
        fill.by = "ident") +
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) +
  stat_summary(fun = median,
        geom = "crossbar",
        width = 0.3,
        size = 0.2,
        position = position_dodge(width = 0.5))

```

## Clusters - subclass level2

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=5, fig.width=10}

Idents(SO) <- "subclass.l2"
sub.l2_levels <- c("PT-S1-a",	"PT-S1-b",	"PT-S1-c",	"PT-S2-a",	"PT-S2-b",	"PT-S2-c",	"PT-S2-d",	"Pericyte-a",	"Pericyte-b",	"Podocyte",	"Endo-a",	"Endo-b",	"Endo-c",	"TAL-1",	"TAL-2",	"Macula Densa",	"DCT1",	"DCT2",	"CNT-1",	"CNT-2",	"PC",	"IC-A",	"IC-B",	"Peri-CNT",	"PT-Peri",	"PTS1-DCT",	"PTS1-Endo",	"PTS1-TAL",	"Endo-TAL", "Endo-DCT",	"Immune-PT-CNT",	"X-a",	"X-b",	"X-c",	"X-d",	"X-e",	"X-f")
Idents(SO) <- factor(x = Idents(SO), levels = sub.l2_levels)

DimPlot(SO, 
        reduction = "umap", 
        label = TRUE,
        repel = TRUE,
        pt.size = 0.5) +
  ggtitle("Subclass level2")
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=5, fig.width=12}

DotPlot(SO,
        features = markers.to.plot2,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.8)) +
  coord_flip() +
  ggtitle("Subclass level2")

VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0,
        fill.by = "ident") +
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) +
  stat_summary(fun = median,
        geom = "crossbar",
        width = 0.3,
        size = 0.2,
        position = position_dodge(width = 0.5))

```

## Create a proportion table - class

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=4, fig.width=12}

Idents(SO) <- "class"
Idents(SO) <- factor(x = Idents(SO), levels = class_levels)

t1 <- table(Idents(SO), SO$Treatment)[, c("Veh", "AngII")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Treatment', 'Frequency')

ggplot(t2,
       aes(fill=Cell_type, x=Frequency, y=Treatment)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic() +
  theme(legend.position = "bottom")

```

## Create a proportion table - subclass.l1

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=4, fig.width=12}

Idents(SO) <- "subclass.l1"
Idents(SO) <- factor(x = Idents(SO), levels = sub.l1_levels)

t1 <- table(Idents(SO), SO$Treatment)[, c("Veh", "AngII")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Treatment', 'Frequency')

ggplot(t2,
       aes(fill=Cell_type, x=Frequency, y=Treatment)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic() +
  theme(legend.position = "bottom")

```

## Create a proportion table - subclass.l2

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=4, fig.width=12}

Idents(SO) <- "subclass.l2"
Idents(SO) <- factor(x = Idents(SO), levels = sub.l2_levels)

t1 <- table(Idents(SO), SO$Treatment)[, c("Veh", "AngII")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Treatment', 'Frequency')

ggplot(t2,
       aes(fill=Cell_type, x=Frequency, y=Treatment)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic() +
  theme(legend.position = "bottom")

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
sessionInfo()
```
