---
title: "PT ACE2 KO PT Pathways"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: false
    number_sections: no
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```


```{r load new packages, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}


if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}


if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("ggupset")) {install.packages("ggupset"); require("ggupset")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}

if (!require("AnnotationHub")) {BiocManager::install('AnnotationHub'); require("AnnotationHub")}
if (!require("org.Mm.eg.db")) {BiocManager::install('org.Mm.eg.db'); require("org.Mm.eg.db")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("stats")) {install.packages("stats"); require("stats")}

library(stats)

#install.packages("kableExtra") # Install the package
library(kableExtra)           # Load the package

#install.packages("scCustomize")
library(scCustomize)

if(!require("magick")) {install.packages("magick",repos = "http://cran.us.r-project.org"); require("magick")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")} # volcano plot


library("Seurat")
library("ggplot2")
library("cowplot")
here()

SO <- readRDS(here("Temp Outputs", "SO_jwn101823.rds"))
```

# Load SO cluster-named dataset

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
Idents(SO)<-"class.jwn2"
DimPlot(SO, reduction = "umap", group.by= "Treatment")
DimPlot(SO)
DimPlot(SO, group.by = "class")

```

# Comparisons

Comp1 =  KO Saline vs. Control Saline (higher in KO Saline positive)
Comp2  = KO AngII vs. Control AngII (higher in KO AngII positive)
Comp3  = Control AngII vs. Control Saline (higher in Control AngII positive)

## KO Saline vs. Control Saline DEGs (Comp1)

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

AllPT_Comp1 <- FindMarkers(SO, ident.1 = "PT ACE2 KO Saline", ident.2 = "Control Saline", group.by = "Experimental_Groups", subset.ident = c("PTS1", "PTS2", "PTS3"), logfc.threshold = .25, min.pct = 0.1)
AllPT_Comp1 <- AllPT_Comp1 %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
AllPT_Comp1 <- AllPT_Comp1[order(AllPT_Comp1[,"avg_log2FC"], decreasing = TRUE),]

AllPT_Comp1

write.xlsx(AllPT_Comp1, file = here("KOvCtrl_Saline_All.xlsx"))

EnhancedVolcano(AllPT_Comp1,
                lab = AllPT_Comp1$gene, 
                title = ("KO-Saline vs. Ctrl-Saline"),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = TRUE, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 5, 
                legendIconSize = 0, 
                drawConnectors = TRUE, 
                xlim = c(-2, 3),
                widthConnectors = 0.75) + 
  theme(plot.title = element_text(hjust = 0.5))



  EnhancedVolcano(AllPT_Comp1,
    lab = AllPT_Comp1$gene,
    x = 'avg_log2FC',
    y = 'p_val_adj',
    selectLab = c('Ace2','Lrp2', "Slc6a19", "Slc6a15", "Slc6a18"),
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 10e-14,
    FCcutoff = 0.25,
    pointSize = 1.0,
    labSize = 5.0,
    colAlpha = 1,
    legendPosition = 'right',
    legendLabSize = 14,
    legendIconSize = 5.0)
  
  
    EnhancedVolcano(AllPT_Comp2,
    lab = AllPT_Comp2$gene,
    x = 'avg_log2FC',
    y = 'p_val_adj',
    selectLab = c('Ace2','Lrp2', "Slc6a19", "Slc6a15", "Slc6a18"),
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 10e-14,
    FCcutoff = 0.25,
    pointSize = 1.0,
    labSize = 5.0,
    colAlpha = 1,
    legendPosition = 'right',
    legendLabSize = 14,
    legendIconSize = 5.0)
```

### KO Saline vs. Control Saline Up

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

markers <- AllPT_Comp1

head(markers, n = 50)

#markers <- markers %>% rownames_to_column(var="SYMBOL")

markers <- markers %>% rename("gene" = "SYMBOL")

#head(markers, n = 50)

ENTREZ_list <- bitr(geneID = markers$SYMBOL,  #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
pos.markers

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0.3219]
length(pos.ranks)

# print the number of values in pos.ranks



pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "all",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go, split="ONTOLOGY") +
        facet_grid(ONTOLOGY~., scale="free") +          #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways") +
        theme(plot.title = element_text(hjust = 0.5))

dotplot(pos_go, split="ONTOLOGY") +
  facet_grid(ONTOLOGY~., scale="free") +
  ggtitle("KO Saline vs. Control Saline Up Pathways") +
  theme_bw() +  # You can use a different theme if needed
  theme(
    plot.title = element_text(hjust = 0.5, size = 12),  # Adjust title font size
    axis.text = element_text(size = 9)  # Adjust axis text font size
  )


Comp1pos_go <- pos_go

genes <- Comp1pos_go@result$GeneID

# Replace 'GO:XXXXXXX' with your specific GO term of interest
go_term <- 'GO:0006119'

# Select the geneID column for the specified GO term
gene_ids <- Comp1pos_go$geneID[Comp1pos_go$ID == go_term]

# Print or use the selected GeneIDs
print(gene_ids)




write.xlsx(Comp1pos_go, file = here("KOvCtrl_Saline_up.xlsx"))


#GO-Biological Processes
pos_go.BP <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(pos_go.BP, showCategory=20)

dotplot(pos_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Biological Processes") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(pos_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Bioological Functions") +
        theme(plot.title = element_text(hjust = 0.5))


#GO - Cellular components
pos_go.CC <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="CC",
                readable=TRUE)

barplot(pos_go.CC, showCategory=20)



dotplot(pos_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(pos_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))

#GO- Molecular Function
pos_go.MF <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="MF",
                readable=TRUE)

barplot(pos_go.MF, showCategory=20) 

dotplot(pos_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Molecular Function") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(pos_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Molecular Functions") +
        theme(plot.title = element_text(hjust = 0.5))




```

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}
#Examine gene expression by PT segment & Experimental group
genes <- unlist(head((pos.markers$SYMBOL), n= 10))

for (i in genes) {
  f1 <- VlnPlot(SO, i, idents = c("PTS1", "PTS2", "PTS3"), split.by = "Experimental_Groups")
  print(f1)
  }

VlnPlot(SO, "Slc9a3", idents = c("PTS1", "PTS2", "PTS3"), split.by = "Experimental_Groups")
VlnPlot(SO, "Col4a4", idents = c("PTS1", "PTS2", "PTS3"), split.by = "Experimental_Groups")
```


### KO Saline vs. Control Saline Down

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

markers <- AllPT_Comp1

head(markers, n = 50)

#markers <- markers %>% rownames_to_column(var="SYMBOL")

markers <- markers %>% rename("gene" = "SYMBOL")

#head(markers, n = 50)

ENTREZ_list <- bitr(geneID = markers$SYMBOL,  #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

neg.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
neg.markers

neg.ranks <- neg.markers$ENTREZID[abs(neg.markers$avg_log2FC) > 0.3219]
length(neg.ranks)

neg_go <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

neg_go

dotplot(neg_go, split="ONTOLOGY") + facet_grid(ONTOLOGY~., scale="free") + ggtitle("KO Saline vs. Control Saline Downregulated Pathways") + theme(plot.title = element_text(hjust = 0.5))




dotplot(neg_go, split="ONTOLOGY") +
  facet_grid(ONTOLOGY~., scale="free") +
  ggtitle("KO Saline vs. Control Saline Down Pathways") +
  theme_bw() +  # You can use a different theme if needed
  theme(
    plot.title = element_text(hjust = 0.5, size = 12),  # Adjust title font size
    axis.text = element_text(size = 8.5
                             )  # Adjust axis text font size
  )



Comp1neg_go <- neg_go

write.xlsx(Comp1neg_go, file = here("KOvCtrl_Saline_down.xlsx"))


neg.markers.ENTREZID <- neg.markers %>%
  dplyr::select(ENTREZID, avg_log2FC) %>%
  column_to_rownames(var = "ENTREZID")

head(neg.markers.ENTREZID)

neg.geneList <- neg.markers.ENTREZID$avg_log2FC
names(neg.geneList) <- rownames(neg.markers.ENTREZID)

head(neg.geneList)


#GO-Biological Processes
neg_go.BP <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(neg_go.BP, showCategory=20)

dotplot(neg_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Biological Processes") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(neg_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Bioological Functions") +
        theme(plot.title = element_text(hjust = 0.5))

#GO - Cellular components
neg_go.CC <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="CC",
                readable=TRUE)

barplot(neg_go.CC, showCategory=20)



dotplot(neg_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(neg_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))

#GO- Molecular Function
neg_go.MF <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="MF",
                readable=TRUE)

barplot(neg_go.MF, showCategory=20) 

dotplot(neg_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Molecular Function") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(neg_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Molecular Function") +
        theme(plot.title = element_text(hjust = 0.5))

#Upset plot

upsetplot(neg_go)
```

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

genes <- unlist(head((neg.markers$SYMBOL), n= 10))

for (i in genes) {
  f1 <- VlnPlot(SO, i, idents = c("PTS1", "PTS2", "PTS3"), split.by = "Experimental_Groups")
  print(f1)
  }


VlnPlot(SO, "Slc6a15", idents = c("PTS1", "PTS2", "PTS3"), split.by = "Experimental_Groups")

VlnPlot(SO, "Cltrn", idents = c("PTS1", "PTS2", "PTS3"), split.by = "Experimental_Groups")

```


## KO AngII vs. Control AngII DEGs (Comp2)

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

AllPT_Comp2 <- FindMarkers(SO, ident.1 = "PT ACE2 KO AngII", ident.2 = "Control AngII", group.by = "Experimental_Groups", subset.ident = c("PTS1", "PTS2", "PTS3"), logfc.threshold = .25, min.pct = 0.1)
AllPT_Comp2 <- AllPT_Comp2 %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
AllPT_Comp2 <- AllPT_Comp2[order(AllPT_Comp2[,"avg_log2FC"], decreasing = TRUE),]

AllPT_Comp2

write.xlsx(AllPT_Comp2, file = here("KOvCtrl_AngII_All.xlsx"))

EnhancedVolcano(AllPT_Comp2,
                lab = AllPT_Comp2$gene, 
                title = ("KO-AngII vs. Ctrl-AngII"),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = NULL, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 5, 
                legendIconSize = 0, 
                drawConnectors = TRUE, 
                xlim = c(-2, 3),
                widthConnectors = 0.75) + 
  theme(plot.title = element_text(hjust = 0.5))

```


### KO AngII vs. Control AngII Up

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

markers <- AllPT_Comp2

head(markers, n = 50)

#markers <- markers %>% rownames_to_column(var="SYMBOL")

markers <- markers %>% rename("gene" = "SYMBOL")

#head(markers, n = 50)

ENTREZ_list <- bitr(geneID = markers$SYMBOL,  #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")



# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0.3219]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "all",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go, split="ONTOLOGY") +
        facet_grid(ONTOLOGY~., scale="free") +          #ggplot2 functions
        ggtitle("KO AngII vs. Control AngII Upregulated Pathways") +
        theme(plot.title = element_text(hjust = 0.5))

dotplot(neg_go, split="ONTOLOGY") +
  facet_grid(ONTOLOGY~., scale="free") +
  ggtitle("KO-AngII vs. Control-AngII Upregulated Pathways") +
  theme_bw() +  # You can use a different theme if needed
  theme(
    plot.title = element_text(hjust = 0.5, size = 12),  # Adjust title font size
    axis.text = element_text(size = 8.5
                             )  # Adjust axis text font size
  )


Comp2pos_go <- pos_go


write.xlsx(Comp2pos_go, file = here("KOvCtrl_AngII_up.xlsx"))


#GO-Biological Processes
pos_go.BP <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(pos_go.BP, showCategory=20)

dotplot(pos_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Biological Processes") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(pos_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Bioological Functions") +
        theme(plot.title = element_text(hjust = 0.5))


goplot(neg_go.BP)


#GO - Cellular components
pos_go.CC <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="CC",
                readable=TRUE)

#barplot(pos_go.CC, showCategory=20) #error when running code



#dotplot(pos_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))

#heatplot(pos_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))
        
        
        

goplot(neg_go.CC)

#GO- Molecular Function
pos_go.MF <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="MF",
                readable=TRUE)

barplot(pos_go.MF, showCategory=20) 

dotplot(pos_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Molecular Function") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(pos_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Molecular Functions") +
        theme(plot.title = element_text(hjust = 0.5))



goplot(neg_go.MF)
```


```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

genes <- unlist(head((pos.markers$SYMBOL), n= 10))

for (i in genes) {
  f1 <- VlnPlot(SO, i, idents = c("PTS1", "PTS2", "PTS3"), split.by = "Experimental_Groups")
  print(f1)
  }

Idents(SO)<-"class"

VlnPlot(SO, "Slc6a18", idents = c("PT"), split.by = "Experimental_Groups")

```

### KO AngII vs. Control AngII Down

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

markers <- AllPT_Comp2

head(markers, n = 50)

#markers <- markers %>% rownames_to_column(var="SYMBOL")

markers <- markers %>% rename("gene" = "SYMBOL")

#head(markers, n = 50)

ENTREZ_list <- bitr(geneID = markers$SYMBOL,  #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

neg.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(neg.markers, n = 50)

neg.ranks <- neg.markers$ENTREZID[abs(neg.markers$avg_log2FC) > 0.3219]
#head(neg.ranks)

neg_go <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

neg_go

dotplot(neg_go, split="ONTOLOGY") + facet_grid(ONTOLOGY~., scale="free") + ggtitle("KO AngII vs. Control AngII Downregulated Pathways") + theme(plot.title = element_text(hjust = 0.5))



dotplot(neg_go, split="ONTOLOGY") +
  facet_grid(ONTOLOGY~., scale="free") +
  ggtitle("KO AngII vs. Control AngII Down Pathways") +
  theme_bw() +  # You can use a different theme if needed
  theme(
    plot.title = element_text(hjust = 0.5, size = 12),  # Adjust title font size
    axis.text = element_text(size = 8.5
                             )  # Adjust axis text font size
  )



Comp2neg_go <- neg_go


write.xlsx(Comp2neg_go, file = here("KOvCtrl_AngII_down.xlsx"))


neg.markers.ENTREZID <- neg.markers %>%
  dplyr::select(ENTREZID, avg_log2FC) %>%
  column_to_rownames(var = "ENTREZID")

head(neg.markers.ENTREZID)

neg.geneList <- neg.markers.ENTREZID$avg_log2FC
names(neg.geneList) <- rownames(neg.markers.ENTREZID)

head(neg.geneList)


#GO-Biological Processes
neg_go.BP <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(neg_go.BP, showCategory=20)

dotplot(neg_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Biological Processes") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(neg_go.BP, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Bioological Functions") +
        theme(plot.title = element_text(hjust = 0.5))

#GO - Cellular components
neg_go.CC <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="CC",
                readable=TRUE)

barplot(neg_go.CC, showCategory=20)



dotplot(neg_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(neg_go.CC, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Cellular Components") +
        theme(plot.title = element_text(hjust = 0.5))

#GO- Molecular Function
neg_go.MF <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="MF",
                readable=TRUE)

barplot(neg_go.MF, showCategory=20) 

dotplot(neg_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Up Pathways", 
                 subtitle = "GO-Molecular Function") +
        theme(plot.title = element_text(hjust = 0.5))

heatplot(neg_go.MF, showCategory=15) +
                #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Down Pathways", 
                 subtitle = "GO-Molecular Function") +
        theme(plot.title = element_text(hjust = 0.5))

#Upset plot

upsetplot(neg_go)

```

## Control AngII vs. Control Saline DEGs (Comp3)

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

AllPT_Comp3 <- FindMarkers(SO, ident.1 = "Control AngII", ident.2 = "Control Saline", group.by = "Experimental_Groups", subset.ident = c("PTS1", "PTS2", "PTS3"), logfc.threshold = .25)
AllPT_Comp3 <- AllPT_Comp3 %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
AllPT_Comp3 <- AllPT_Comp3[order(AllPT_Comp3[,"avg_log2FC"], decreasing = TRUE),]

AllPT_Comp3

EnhancedVolcano(AllPT_Comp3,
                lab = AllPT_Comp3$gene, 
                title = ("AngII vs. Saline Control"),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = NULL, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 5, 
                legendIconSize = 0, 
                drawConnectors = TRUE, 
                xlim = c(-2, 3),
                widthConnectors = 0.75) + 
  theme(plot.title = element_text(hjust = 0.5))

```


### Control AngII vs. Control Saline Up

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T, fig.height=15}

markers <- AllPT_Comp3

head(markers, n = 50)

#markers <- markers %>% rownames_to_column(var="SYMBOL")

markers <- markers %>% rename("gene" = "SYMBOL")

#head(markers, n = 50)

ENTREZ_list <- bitr(geneID = markers$SYMBOL,  #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0.3219]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "all",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go, split="ONTOLOGY") +
        facet_grid(ONTOLOGY~., scale="free") +          #ggplot2 functions
        ggtitle("KO Saline vs. Control Saline Upregulated Pathways") +
        theme(plot.title = element_text(hjust = 0.5))


Comp3pos_go <- pos_go

```

### KO AngII vs. Control AngII Down

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T, fig.height=15}

markers <- AllPT_Comp3

head(markers, n = 50)

#markers <- markers %>% rownames_to_column(var="SYMBOL")

markers <- markers %>% rename("gene" = "SYMBOL")

#head(markers, n = 50)

ENTREZ_list <- bitr(geneID = markers$SYMBOL,  #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

neg.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(neg.markers, n = 50)

neg.ranks <- neg.markers$ENTREZID[abs(neg.markers$avg_log2FC) > 0.3219]
head(neg.ranks)

neg_go <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

neg_go

dotplot(neg_go, split="ONTOLOGY") + facet_grid(ONTOLOGY~., scale="free") + ggtitle("Control AngII vs. Control Saline Downregulated Pathways") + theme(plot.title = element_text(hjust = 0.5))

Comp3neg_go <- neg_go

```

# Save DEG Lists and GO files

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T, fig.height=15}

save(AllPT_Comp1, Comp1pos_go, Comp1neg_go, AllPT_Comp2, Comp2pos_go, Comp2neg_go, AllPT_Comp3, Comp3pos_go, Comp3neg_go, file = here("Outputs", "Pathways_data.rdata"))

```

# Session Info

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

sessionInfo()

```



