---
title: "PT ACE2 KO Merge and Integration "
author: "Jonathan Nelson and Jackie Emathinger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: 
      collapsed: true
      smooth_scroll: true
    number_sections: false
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=T, error=FALSE, message=FALSE, warning=FALSE}

library("dplyr")
library("Seurat")
library("knitr")
library("ggplot2")
library("BiocManager")
library("here")
#BiocManager::install("EnhancedVolcano")
library("EnhancedVolcano") #volcano plot
#BiocManager::install('DESeq2') #for DEG
library("DESeq2")
library("tidyverse") #tidy up data

if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")} # for color brewer
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")} # for data normalization
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")} # for data normalization, sctransform
if (!require("cowplot")) {install.packages("cowplot"); require("cowplot")} # for figure layout
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")} # for figure patching
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")} # to save .xlsx files
set.seed(12345)
here()
```

# Starting Place

I'm starting with the allcombined object which hasn't been filtered except for nFeatures and nCounts.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- readRDS(here("Temp Outputs", "allcombined.rds"))

SO

head(SO@meta.data)

DimPlot(SO) + ggtitle("JME Cluster Identification")

```

# Round 1: Identify and Remove Doublets

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Idents(SO) <- SO@meta.data$seurat_clusters

DimPlot(SO, label = T) + ggtitle("Res = 1 clusters")

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 12, fig.height= 8}

markers.to.plot1 <- c("Lrp2",         # PT
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Havcr1",       # Injured PT
                      "Epha7",        # dTL
                      "Cryab",        # dTL
                      "Cdh13",        # dTL1
                      "Slc14a2",      # dTL2
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1,
                      "Cldn10",       # TAL
                      "Cldn16",       # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2, CNT
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs1",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Emcn",         # Glom Endo
                      "Kdr",          # Capillary Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Piezo2",       # Mesangial
                      "Acta2",        # Mural
                      "Ptprc",        # Immune
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b",        # Uro
                      "Top2a"         # Proliferation
)
                      
DotPlot(SO,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

```

## Initial Cluster Naming

I found chatGPT really helpful here, editing the code to make it easy to use. 

Step 1: copy and paste the chunck of code below.
Step 2: "can you extend this code to 51"
Step 3: "Synthesize with this list" + copy and paste Excel annotations with numbers

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO@meta.data <- SO@meta.data %>% mutate(subclass.jwn = dplyr::case_when(
seurat_clusters == 0 ~ "EC-a",
seurat_clusters == 1 ~ "PTS2-a",
seurat_clusters == 2 ~ "PTS2-b",
seurat_clusters == 3 ~ "PTS1-a",
seurat_clusters == 4 ~ "PTS3-a",
seurat_clusters == 5 ~ "TAL-Cldn10",
seurat_clusters == 6 ~ "PTS1-PTS2",
seurat_clusters == 7 ~ "DCT1",
seurat_clusters == 8 ~ "TAL-Cldn16",
seurat_clusters == 9 ~ "CNT",
seurat_clusters == 10 ~ "Fib-a",
seurat_clusters == 11 ~ "PTS1-b",
seurat_clusters == 12 ~ "PC-a",
seurat_clusters == 13 ~ "EC-b",
seurat_clusters == 14 ~ "PTS3-b",
seurat_clusters == 15 ~ "PC-b",
seurat_clusters == 16 ~ "IC-A",
seurat_clusters == 17 ~ "DCT2",
seurat_clusters == 18 ~ "IC-B",
seurat_clusters == 19 ~ "Podo",
seurat_clusters == 20 ~ "EC-c",
seurat_clusters == 21 ~ "dTL2",
seurat_clusters == 22 ~ "Immune",
seurat_clusters == 23 ~ "aTL",
seurat_clusters == 24 ~ "dTL1-a",
seurat_clusters == 25 ~ "PT-inj",
seurat_clusters == 26 ~ "Uro",
seurat_clusters == 27 ~ "PT-EC",
seurat_clusters == 28 ~ "dTL1-b",
seurat_clusters == 29 ~ "Contractile",
seurat_clusters == 30 ~ "PTS2-c",
seurat_clusters == 31 ~ "Fib-b",
seurat_clusters == 32 ~ "PEC",
seurat_clusters == 33 ~ "MD",
seurat_clusters == 34 ~ "PT-TAL-a",
seurat_clusters == 35 ~ "Fib-TAL",
seurat_clusters == 36 ~ "PC-weak",
seurat_clusters == 37 ~ "Mes",
seurat_clusters == 38 ~ "PT-TAL-b",
seurat_clusters == 39 ~ "PT-TAL-c",
seurat_clusters == 40 ~ "TAL-PC",
seurat_clusters == 41 ~ "TAL-DCT",
seurat_clusters == 42 ~ "X"
      ))

```

## Compare JME to JWN Cluster Labeling

90% Similar with only minor changes at the end. I think some of the clusters I may have called doublets that JME did not. JME identified Urothelial Cells (great spot, which I missed in my initial naming.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

selected_columns <- SO@meta.data[, c("seurat_clusters", "subclass.l2", "subclass.jwn")]

unique_rows <- selected_columns %>%
  distinct(seurat_clusters, .keep_all = TRUE) %>%
  arrange(seurat_clusters)

unique_rows

```

## Rename Clusters and Filter Out Doublets

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO@meta.data <- SO@meta.data %>% mutate(class.jwn = dplyr::case_when(
seurat_clusters == 0 ~ "EC",
seurat_clusters == 1 ~ "PTS2",
seurat_clusters == 2 ~ "PTS2",
seurat_clusters == 3 ~ "PTS1",
seurat_clusters == 4 ~ "PTS3",
seurat_clusters == 5 ~ "TAL",
seurat_clusters == 6 ~ "PTS1",
seurat_clusters == 7 ~ "DCT1",
seurat_clusters == 8 ~ "TAL",
seurat_clusters == 9 ~ "CNT",
seurat_clusters == 10 ~ "Fib",
seurat_clusters == 11 ~ "PTS1",
seurat_clusters == 12 ~ "PC",
seurat_clusters == 13 ~ "EC",
seurat_clusters == 14 ~ "PTS3",
seurat_clusters == 15 ~ "PC",
seurat_clusters == 16 ~ "IC-A",
seurat_clusters == 17 ~ "DCT2",
seurat_clusters == 18 ~ "IC-B",
seurat_clusters == 19 ~ "Podo",
seurat_clusters == 20 ~ "EC",
seurat_clusters == 21 ~ "dTL",
seurat_clusters == 22 ~ "Immune",
seurat_clusters == 23 ~ "dTL",
seurat_clusters == 24 ~ "aTL",
seurat_clusters == 25 ~ "PT-inj",
seurat_clusters == 26 ~ "Uro",
seurat_clusters == 27 ~ "Doublet",
seurat_clusters == 28 ~ "dTL",
seurat_clusters == 29 ~ "Contractile",
seurat_clusters == 30 ~ "PTS2",
seurat_clusters == 31 ~ "Fib",
seurat_clusters == 32 ~ "PEC",
seurat_clusters == 33 ~ "MD",
seurat_clusters == 34 ~ "Doublet",
seurat_clusters == 35 ~ "Doublet",
seurat_clusters == 36 ~ "Doublet",
seurat_clusters == 37 ~ "Mes",
seurat_clusters == 38 ~ "Doublet",
seurat_clusters == 39 ~ "Doublet",
seurat_clusters == 40 ~ "Doublet",
seurat_clusters == 41 ~ "Doublet",
seurat_clusters == 42 ~ "Doublet"
   ))

Idents(SO) <- SO@meta.data$class.jwn

DimPlot(SO) + ggtitle("With Doublets")

SO2 <- subset(SO, idents = "Doublet", invert = T)

DimPlot(SO2, label = T) + ggtitle("Doublets Removed")

```

# Round 2: Re-graph and Super-cluster Without Doublets

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO2 <- SCTransform(SO2, verbose = F)
SO2 <- RunPCA(SO2, verbose = F)
SO2 <- RunUMAP(SO2, dims = 1:30, verbose = F)
SO2 <- FindNeighbors(SO2, dims = 1:30, verbose = F)
SO2 <- FindClusters(SO2, resolution = 2, verbose = T)

DimPlot(SO2, label = T, group.by = "class.jwn") + ggtitle("Previous Cluster Labels")
DimPlot(SO2, label = T) + ggtitle("Super Clustering")

```

## Dotplot to identify cell populations 

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 16, fig.height=10}

DotPlot(SO2,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```

## Re-name cell identities

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 10, fig.width = 10}

SO2@meta.data <- SO2@meta.data %>% mutate(subclass.jwn2 = dplyr::case_when(
seurat_clusters == 0 ~ "TAL-Cldn16",
seurat_clusters == 1 ~ "EC-a",
seurat_clusters == 2 ~ "TAL-Cldn10",
seurat_clusters == 3 ~ "Fib-a",
seurat_clusters == 4 ~ "PC",
seurat_clusters == 5 ~ "PTS2-a",
seurat_clusters == 6 ~ "EC-Cap",
seurat_clusters == 7 ~ "PTS1-a",
seurat_clusters == 8 ~ "PTS1-PTS2",
seurat_clusters == 9 ~ "PTS2-b",
seurat_clusters == 10 ~ "PTS2-b",
seurat_clusters == 11 ~ "DCT1",
seurat_clusters == 12 ~ "PTS1-b",
seurat_clusters == 13 ~ "PTS3-a",
seurat_clusters == 14 ~ "PTS2-c",
seurat_clusters == 15 ~ "PTS1-c",
seurat_clusters == 16 ~ "TAL-Cldn10",
seurat_clusters == 17 ~ "EC-Glom",
seurat_clusters == 18 ~ "PTS2-d",
seurat_clusters == 19 ~ "DCT1",
seurat_clusters == 20 ~ "PTS3-b",
seurat_clusters == 21 ~ "PTS3-c",
seurat_clusters == 22 ~ "CNT-PC",
seurat_clusters == 23 ~ "dTL",
seurat_clusters == 24 ~ "PTS1-d",
seurat_clusters == 25 ~ "PTS1-e",
seurat_clusters == 26 ~ "IC-A",
seurat_clusters == 27 ~ "DCT2",
seurat_clusters == 28 ~ "IC-B",
seurat_clusters == 29 ~ "Podo",
seurat_clusters == 30 ~ "CNT",
seurat_clusters == 31 ~ "EC-b",
seurat_clusters == 32 ~ "dTL-a",
seurat_clusters == 33 ~ "PC-a",
seurat_clusters == 34 ~ "PEC",
seurat_clusters == 35 ~ "PC-b",
seurat_clusters == 36 ~ "CNT",
seurat_clusters == 37 ~ "Uro",
seurat_clusters == 38 ~ "Macro",
seurat_clusters == 39 ~ "PTS3-d",
seurat_clusters == 40 ~ "Fib-b",
seurat_clusters == 41 ~ "dTL-c",
seurat_clusters == 42 ~ "PTS2-d",
seurat_clusters == 43 ~ "?",
seurat_clusters == 44 ~ "Contractile",
seurat_clusters == 45 ~ "EC-c",
seurat_clusters == 46 ~ "MD",
seurat_clusters == 47 ~ "PT-TAL",
seurat_clusters == 48 ~ "Mes",
seurat_clusters == 49 ~ "B & T",
seurat_clusters == 50 ~ "PT-EC",
seurat_clusters == 51 ~ "TAL-EC"
))

DimPlot(SO2, group.by = "subclass.jwn2", label = T) + ggtitle("subclass.jwn2 labels")

SO2@meta.data <- SO2@meta.data %>% mutate(class.jwn2 = dplyr::case_when(
seurat_clusters == 0 ~ "TAL",
seurat_clusters == 1 ~ "EC",
seurat_clusters == 2 ~ "TAL",
seurat_clusters == 3 ~ "Fib",
seurat_clusters == 4 ~ "PC",
seurat_clusters == 5 ~ "PTS2",
seurat_clusters == 6 ~ "EC",
seurat_clusters == 7 ~ "PTS1",
seurat_clusters == 8 ~ "PTS1",
seurat_clusters == 9 ~ "PTS2",
seurat_clusters == 10 ~ "PTS2",
seurat_clusters == 11 ~ "DCT1",
seurat_clusters == 12 ~ "PTS1",
seurat_clusters == 13 ~ "PTS3",
seurat_clusters == 14 ~ "PTS2",
seurat_clusters == 15 ~ "PTS1",
seurat_clusters == 16 ~ "TAL",
seurat_clusters == 17 ~ "EC",
seurat_clusters == 18 ~ "PTS2",
seurat_clusters == 19 ~ "DCT1",
seurat_clusters == 20 ~ "PTS3",
seurat_clusters == 21 ~ "PTS3",
seurat_clusters == 22 ~ "CNT",
seurat_clusters == 23 ~ "dTL",
seurat_clusters == 24 ~ "PTS1",
seurat_clusters == 25 ~ "PTS1",
seurat_clusters == 26 ~ "ICA",
seurat_clusters == 27 ~ "DCT2",
seurat_clusters == 28 ~ "ICB",
seurat_clusters == 29 ~ "Podo",
seurat_clusters == 30 ~ "CNT",
seurat_clusters == 31 ~ "EC",
seurat_clusters == 32 ~ "dTL",
seurat_clusters == 33 ~ "PC",
seurat_clusters == 34 ~ "PEC",
seurat_clusters == 35 ~ "PC",
seurat_clusters == 36 ~ "CNT",
seurat_clusters == 37 ~ "Uro",
seurat_clusters == 38 ~ "Macro",
seurat_clusters == 39 ~ "PTS3",
seurat_clusters == 40 ~ "Fib",
seurat_clusters == 41 ~ "dTL",
seurat_clusters == 42 ~ "PTS2",
seurat_clusters == 43 ~ "EC",
seurat_clusters == 44 ~ "Contractile",
seurat_clusters == 45 ~ "EC",
seurat_clusters == 46 ~ "MD",
seurat_clusters == 47 ~ "Doublet",
seurat_clusters == 48 ~ "Mes",
seurat_clusters == 49 ~ "Lympho",
seurat_clusters == 50 ~ "Doublet",
seurat_clusters == 51 ~ "Doublet"
))

DimPlot(SO2, group.by = "class.jwn2", label = T) + ggtitle("class.jwn labels")

SO2@meta.data$class.jwn2 <- factor(SO2@meta.data$class.jwn2, levels = c("PTS1", "PTS2", "PTS3", "dTL", "TAL", "MD", "DCT1", "DCT2", "CNT", "PC", "ICA", "ICB", "Podo", "PEC", "EC", "Fib", "Contractile", "Mes", "Macro", "Lympho", "Uro", "Doublet"))

Idents(SO2) <- SO2@meta.data$class.jwn2

SO3 <- subset(SO2, idents = "Doublet", invert = T)

DimPlot(SO3, group.by = "class.jwn2") + ggtitle("After Cluster Leveling")

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 14, fig.height = 10}

DotPlot(SO3,
features = markers.to.plot1,
group.by = "class.jwn2",
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

markers.to.plot.clean <- c(
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Epha7",        # dTL
                      "Slc12a1",      # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs1",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Acta2",        # Mural
                      "Piezo2",       # Mesangial
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b"         # Uro
                      )

DotPlot(SO3,
features = markers.to.plot.clean,
group.by = "class.jwn2",
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

```

# Final DimPlot and DotPlot

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 10, fig.width = 10}

DimPlot(SO3, group.by = "class.jwn2", label = T) + ggtitle("Final Clustering")

DimPlot(SO3, group.by = "class.jwn2") + ggtitle("Final Clustering")

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 14, fig.height = 10}

DotPlot(SO3,
features = markers.to.plot.clean,
group.by = "class.jwn2",
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

## Change Metadata labels

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 14, fig.height = 10}


#SO2 <- readRDS(here("Temp Outputs", "SO_jwn090623.rds"))

SO3@meta.data$Genotype_Treatment <- paste0(SO3@meta.data$Genotype, "_", SO3@meta.data$Treatment)
DimPlot(SO3, reduction = "umap", group.by= "Genotype_Treatment")

# Re-order groups
Genotype_Treatment_levels <- c("Cre -_Saline",  "Cre +_Saline", "Cre -_AngII", "Cre +_AngII")
SO3@meta.data$Genotype_Treatment <- factor(x = SO3@meta.data$Genotype_Treatment, levels = Genotype_Treatment_levels)
DimPlot(SO3, reduction = "umap", group.by="Genotype_Treatment")

SO3@meta.data <- SO3@meta.data %>% mutate(Experimental_Groups = dplyr::case_when(
Genotype_Treatment == "Cre -_Saline" ~ "Control Saline",
Genotype_Treatment == "Cre +_Saline" ~ "PT ACE2 KO Saline",
Genotype_Treatment == "Cre -_AngII" ~ "Control AngII",
Genotype_Treatment == "Cre +_AngII" ~ "PT ACE2 KO AngII",
))

Experimental_Groups_levels <- c("Control Saline",  "PT ACE2 KO Saline", "Control AngII", "PT ACE2 KO AngII")
SO3@meta.data$Experimental_Groups <- factor(x = SO3@meta.data$Experimental_Groups, levels = Experimental_Groups_levels)

DimPlot(SO3, reduction = "umap", group.by="Experimental_Groups")

## Labeling Convention Asked for by JME
## Cre- = Control
## Cre+ = PT ACE2 KO
## PT ACE2 KO Saline

```


# Save Object

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(SO3, here("Temp Outputs", "SO_jwn101823.rds"))

```

# Session Info

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```
