---
title: "Supplemental Figures"
author: "Jackie Emathinger & Jonathan Nelson"
date: "2023-12-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries & Seurat Object

```{r load new packages, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}


if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}


if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("ggupset")) {install.packages("ggupset"); require("ggupset")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}

if (!require("AnnotationHub")) {BiocManager::install('AnnotationHub'); require("AnnotationHub")}
if (!require("org.Mm.eg.db")) {BiocManager::install('org.Mm.eg.db'); require("org.Mm.eg.db")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("stats")) {install.packages("stats"); require("stats")}

library(stats)

#install.packages("kableExtra") # Install the package
library(kableExtra)           # Load the package

#install.packages("scCustomize")
library(scCustomize)

if(!require("magick")) {install.packages("magick",repos = "http://cran.us.r-project.org"); require("magick")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")} # volcano plot


library("Seurat")
library("ggplot2")
library("cowplot")

library("stringr")
here()

SO <- readRDS(here("GSE253448_PTACE2KO.rds"))
```


## Supplemental Figure 2: Cannonical Marker Expression for all Cell Types
### Supplemental Figure 2A

```{r vlnplot, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=6, fig.width=16}
fS2 <- DimPlot(SO,
              reduction = "umap",
              label = TRUE,
              pt.size = 0,
              label.size = 6) +
  NoLegend()

markers.to.plot1 <- c(
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Epha7",        # dTL
                      "Slc12a1",      # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs1",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Acta2",        # Mural
                      "Piezo2",       # Mesangial
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b"         # Uro
                      )

fS2A <- VlnPlot(SO, 
        features = markers.to.plot1,
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0,
        fill.by = "ident") +
  theme(legend.position = "none", 
              plot.title = element_text(hjust = 0.5),
              axis.text.x = element_text(angle = 0, hjust = .5),
              axis.title.x = element_blank()) +  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  stat_summary(fun = median,
               geom = "crossbar",
               width = 0.1,
               size = 0.1,
               position = position_dodge(width = 0.5))
fS2A




ggdraw() +
  draw_plot(fS2, x = 0, y = 0, width = .3, height = 1) +
  draw_plot(fS2A, x = .3, y = 0, width = .7, height = 1)



```

### Supplemental Figure 2B

```{r featureplot, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 6, fig.width = 8}

fS2B <- FeaturePlot(SO,
             features = markers.to.plot1,
             cols = c("lightgrey", "royal blue"),
             combine = F)

for(i in 1:length(fS2B)) {
  fS2B[[i]] <- fS2B[[i]] + NoLegend() + NoAxes()
}
cowplot::plot_grid(plotlist = fS2B)


```


## Supplemental Figure 3: Quality Control of All Nuclei
### Supplemental Figure 3, A-B

```{r QC, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=5, fig.width=12}

fS3A <- VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        fill.by = "ident",
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0) +
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 0, hjust = .5),
        axis.title.x = element_blank()) +
  stat_summary(fun = median,
        geom = "crossbar",
        width = 0.3,
        size = 0.2,
        position = position_dodge(width = 0.5))
fS3A



fS3B<- FeaturePlot(SO,
             features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
             cols = c("lightgrey", "royal blue"),
             ncol = 3)
fS3B
```
## Supplemental Figure 4: Representation of Cell Types Across Experimental Groups
### Supplemental Figure 4A

```{r exp groups, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 5, fig.width = 16}

f4A <- DimPlot(SO, reduction = "umap", split.by = "Experimental_Groups", ncol = 4)


f4A


```

### Supplemental Figure 4B

```{r proportions, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 5, fig.width = 5}

t1 <- table(Idents(SO), SO$Experimental_Groups)[, c("Control Saline", "Control AngII", "PT ACE2 KO Saline", "PT ACE2 KO AngII")]
t1

prop.t1 <- prop.table(t1, margin = 2)
prop.t1

t2 <- as.data.frame(t1)
t2

colnames(t2) <- c('Cell_type', 'Rep', 'Cell_number')
t2

t3 <- t2 %>%
  pivot_wider(names_from = Cell_type, values_from = Cell_number)
t3

t4 <- t3 %>% 
  add_column(Experimental_Groups = c("Control Saline", "Control AngII", "PT ACE2 KO Saline", "PT ACE2 KO AngII"))
t4

t5 <- t4 %>%
  pivot_longer(c("PTS1", "PTS2", "PTS3", "dTL", "TAL", "MD", "DCT1", "DCT2", "CNT", "PC", "ICA", "ICB", "Podo", "PEC", "EC", "Fib", "Contractile", "Mes", "Macro", "Lympho", "Uro"), names_to ="Cell_type", values_to = "Cell_number")
t5

t6 <- t2 %>% 
  pivot_wider(names_from = Cell_type, values_from = Cell_number) %>%
  add_column(Experimental_Groups = c("Control Saline", "Control AngII", "PT ACE2 KO Saline", "PT ACE2 KO AngII")) %>%
  pivot_longer(c("PTS1", "PTS2", "PTS3", "dTL", "TAL", "MD", "DCT1", "DCT2", "CNT", "PC", "ICA", "ICB", "Podo", "PEC", "EC", "Fib", "Contractile", "Mes", "Macro", "Lympho", "Uro"), names_to ="Cell_type", values_to = "Cell_number")

t5 %>%
  arrange(Cell_number) %>%
  mutate(Cell_type = factor(Cell_type, levels=c("PTS1", "PTS2", "PTS3", "dTL", "TAL", "MD", "DCT1", "DCT2", "CNT", "PC", "ICA", "ICB", "Podo", "PEC", "EC", "Fib", "Contractile", "Mes", "Macro", "Lympho", "Uro"))) %>%
ggplot(
       aes(x=Rep,
           y=Cell_number,
           fill=Cell_type)) +     
  geom_bar(position = "fill",
           stat = "identity",
           colour="black") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Distribution of cell types in experimental groups") +
  theme(plot.title = element_text(hjust = 0)) + 
  xlab('')+ ylab('')

```

## Supplemental Figure 5A: Enrichment of apical plasma membrane and brush border pathways 

```{r brush border, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T, fig.height = 6, fig.width = 10}
SO_PT <- subset(SO,  idents = c("PTS1", "PTS2", "PTS3"))

# Significant pathway enrichment for GO:0005903 Brush Border in PT ACE2 KO kidneys
brush_border<- c("Slc22a12", "Myh10", "Eps8", "Car4", "Aqp1", "Myh14", "Slc11a2", "Lrp2", "Ace2", "Myo1e", "Gna12", "Prom1", "Cltrn", "Enpep", "Lima1", "Abcc2", "Pth1r", "Slc6a19",  "Folr1",  "Shank2", "Slc5a2",  "Amn", "Add3", "Slc34a1", "Slc7a9", "Abcg2",  "Slc2a2",  "Myo1d", "Myo18a", "Slc22a5")


brush_border<- list(brush_border)
brush_border

SO_PT <- AddModuleScore(object = SO_PT, features = brush_border, name = "brush_border")

#Significant pathway enrichment for GO:0016324 Apical Plasma Membrane in PT ACE2 KO groups
apical_membrane<- c("Slc10a2", "Abcb1b", "Slc22a12", "Nrg1", "Cyp4a12a", "Slc44a4", "Slc39a8", "Pld1", "Abcc4", "Abcc6", "Mpdz", "Car4", "Aqp1", "Nox4", "Cyp4a10", "Shroom3", "Slc11a2", "Itpk1", "Stx3", "Lrp2", "Ace2", "Slc16a2", "Slc2a5", "Slc1a1", "Slc12a2", "Msn", "Dpp4", "Cyp4a14", "Prom1", "Slc22a8", "Enpep", "Abcc2", "Pth1r", "Slc7a8", "Slc13a1", "Atp2b1", "Cyba", "Akap7", "Slc6a19", "Folr1", "Car12", "Shank2", "Slc5a2", "Amn", "Slc17a1", "Pard3b", "Erbb3", "Slc34a1", "Slc7a9", "Shroom4", "Abcg2", "Prkaa2", "Slc2a2", "Ctsb", "Akr1a1", "Ctnnb1", "Slc6a8", "Slc22a5")
apical_membrane<- list(apical_membrane)
apical_membrane


SO_PT <- AddModuleScore(object = SO_PT, features = apical_membrane, name = "apical_membrane")


# Significant pathway enrichment for GO:0031253 cell projection membrane
cell_projection <- c("Slc22a12", "Eps8", "Sh3yl1", "Car4", "Efcab7", "Aqp1", "Pde6a", "Iqce", "Slc11a2", "Lrp2", "Plekha1", "Ace2", "Gna12", "Bbs7", "Inpp5k", "Msn", "Prom1", "Cltrn", "Enpep", "Lima1",  "Abcc2", "Pth1r", "Slc7a8", "Gabarapl1", "Atp2b1", "Slc6a19",  "Folr1",  "Spata13", "Ptprj", "Shank2", "Appl2", "Slc5a2", "Amn", "Epb41l3", "Cask", "Slc34a1", "Slc7a9", "Abcg2", "Insr", "Slc22a2",  "Myo1d", "Ctnnb1", "Slc22a5")
cell_projection<- list(cell_projection)
cell_projection

SO_PT <- AddModuleScore(object = SO_PT, features = cell_projection, name = "cell_projection")

#Significant pathway enrichment for GO:0032956: regulation of actin cytoskeleton organisation
regulation_actin <- c("Cit", "Fhod3", "Eps8", "Fmn1", "Nrp1", "Fchsd2", "Evl", "Nox4", "Smad3", "Dixdc1", "Serpinf2", "Arhgap18", "Alms1", "Cgnl1", "Inpp5k", "Limch1", "Id1", "Lima1",    "Cotl1",    "Add3",  "Cyfip2", "Baiap2", "Phldb2", "F11r", "Cd2ap", "Arhgap35", "Tsc1")
regulation_actin<- list(regulation_actin)
regulation_actin

SO_PT <- AddModuleScore(object = SO_PT, features = regulation_actin, name = "regulation_actin")


#Significant pathway enrichment for GO:0098862: cluster of actin-based cell projections
actin_projection <- c("Slc22a12", "Myh10", "Eps8", "Dcdc2a", "Clic5", "Fchsd2", "Car4", "Aqp1", "Myh14", "Slc11a2", "Lrp2", "Ace2", "Myo1e", "Gna12", "Dock4", "Zswim6", "Prom1", "Cltrn", "Enpep", "Lima1",  "Abcc2", "Pth1r",  "Slc6a19",  "Folr1",  "Shank2", "Slc5a2",  "Amn", "Add3", "Slc34a1", "Slc7a9", "Abcg2",  "Slc2a2", "Myo1d", "Myo18a", "Slc22a5")

actin_projection<- list(actin_projection)
actin_projection

SO_PT <- AddModuleScore(object = SO_PT, features = actin_projection, name = "actin_projection")


#Significant enrichment of GO:0110053: regulation of actin filament organization
actin_filament<- c("Cit", "Fhod3", "Eps8", "Fmn1", "Nrp1", "Fchsd2", "Evl", "Nox4", "Smad3", "Serpinf2", "Arhgap18", "Alms1", "Cgnl1", "Inpp5k", "Limch1", "Id1", "Lima1",    "Cotl1",  "Add3", "Cyfip2", "Baiap2", "Phldb2", "F11r", "Arhgap35", "Tsc1")

actin_filament<- list(actin_filament)
actin_filament

SO_PT <- AddModuleScore(object = SO_PT, features = actin_filament, name = "actin_filament")


# Create dotplot to visualize the impact of PT ACE2 loss on PT brush border

PT_brushborder<- DotPlot(SO_PT,
  features = c( "brush_border1", "apical_membrane1", "cell_projection1", "regulation_actin1", "actin_projection1", "actin_filament1"),
  group.by = "Experimental_Groups",
  dot.scale = 8,
  dot.min = 0,
  scale.max = 100,
  scale.min = 0,
  col.min = -2.5,
    cols = c("dodgerblue", "firebrick1"), 
  col.max = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip() +
  scale_x_discrete(labels = c( "brush border", "apical plasma membrane", "cell projection membrane", "regulation of actin cytoskeleton organization", "cluster of actin-based cell projections", "regulation of actin filament organization"))+ 
  xlab('')+ ylab('')

print(PT_brushborder)
```

## Supplemental Figure 6: Transcriptional profile of key sodium transporters
### Supplemental Figure 6A

```{r proximal transporters, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 6, fig.width = 8}

#subset out portions of the nephron that express different transporters
Proximal <- subset(SO,  idents = c("PTS1", "PTS2", "PTS3", "TAL"))


#Create lists of transporter genes within the PT and TAL
Transporters_PT<- c("Slc9a3", "Slc5a2", "Cldn2", "Aqp1", "Atp1a1", "Atp1b1", "Cldn10")

#Create dotplot of transporters in the PT and TAL


f6A<-DotPlot(Proximal,
  features = Transporters_PT,
  group.by = "Experimental_Groups",
  dot.scale = 8,
  dot.min = 0,
  scale.max = 75,
  scale.min = 0,
  col.min = -2.5,
 cols = c("grey85", "blue"), 
  col.max = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title = "Proximal nephron transporters") + 
  coord_flip() + 
  xlab('')+ ylab('')

f6A

```

### Supplemental Figure 6B

```{r distal transporters, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height = 6, fig.width = 8}
#subset out portions of the nephron that express different transporters
Distal <- subset(SO, idents= c("TAL", "MD", "DCT1", "DCT2", "CNT", "ICA", "ICB", "PC", "Uro"))

#Create lists of transporter genes within the distal nephron
Transporters_DT <- c("Atp1a1", "Atp1b1", "Slc12a1", "Slc12a3", "Scnn1a", "Scnn1g", "Cldn7", "Cldn8", "Kcnj1", "Aqp2")

#Create dotplot of transporters in the distal nephron

f6B<-DotPlot(Distal,
  features = Transporters_DT,
  group.by = "Experimental_Groups",
  dot.scale = 8,
  dot.min = 0,
  scale.max = 75,
  scale.min = 0,
  col.min = -2.5,
 cols = c("grey85", "blue"), 
  col.max = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title = "Distal nephron transporters") + 
  coord_flip() + 
  xlab('')+ ylab('')

f6B
```


